---
title: "Ontario Aquatic Ecosystem Classification Biotic Modeling"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

The goal of this project is to use data from [Ontario's Aquatic Ecosystem Classification (AEC)](https://geohub.lio.gov.on.ca/maps/mnrf::aquatic-ecosystem-classification-aec-for-ontario/about) together with [Flowing Waters Information System (FWIS)](https://www.comap.ca/fwis/) to develop predictive models of fish and benthic invertebrate communities across Ontario.

```{r overviewMap, echo=FALSE}
library(tmap)
library(tidyverse)
library(sf)

td<-tempdir()

lu_master<-readRDS(file.path("data","lookups.rds"))

aec_region<-lapply(lu_master$stream_packages,function(x) file.path("data","raw","GIS",x))
names(aec_region)<-sapply(names(aec_region),function(x) file.path("data","raw","GIS",x))

sub_regions<-lapply(names(aec_region),function(aes_nm){
  unzip(aes_nm,exdir=td)
  
  zip_cont<-list.files(file.path(td,gsub(".zip","",basename(aes_nm))),recursive = T,full.names = T)
  sub_regions<-zip_cont[grepl("/w",zip_cont) & grepl("AEC_Class.lyrx",zip_cont)]
  sub_regions<-gsub("_Class\\.lyrx","_Core\\.gdb",sub_regions)
  names(sub_regions)<-sub_regions
  
  sub_regions<-lapply(sub_regions, st_layers)
  
  sub_region_out<-map2(names(sub_regions),sub_regions,function(src,lyr) {
    list(boundary=read_sf(src,lyr$name[grepl("Boundary",lyr$name)]),
         stream=read_sf(src,lyr$name[grepl("Reach",lyr$name)]))
  })
  
  names(sub_region_out)<-sapply(names(sub_regions),function(x) gsub(".gdb","",basename(x)))
  
  return(sub_region_out)
})

boundary<-map(sub_regions,~map(.x,~.x$boundary)) %>% 
  unlist(recursive = F) %>% 
  bind_rows()

sub_region_out<-readRDS(file.path("data","final","int_results.rds"))
matched_points<-map(sub_region_out,~.x$bio_pnt_out)

keep_wu<-sapply(matched_points,function(x) nrow(x)>20)
keep_wu<-names(keep_wu)[keep_wu]
keep_wu<-str_extract(keep_wu,"^...")

matched_points2<-matched_points[str_extract(names(matched_points),"^...") %in% keep_wu]

matched_points2<-map2(matched_points2,names(matched_points2),~mutate(.x,WorkUnitName=.y)) %>% 
  bind_rows()

tm_shape(boundary) +
  tm_polygons() +
  tm_shape(matched_points2) +
  tm_dots() +
  tm_layout(main.title="Fish Sampling Locations Across Ontario") +
  tm_credits(paste(length(unique(matched_points2$SampleEventID)),"Sampling events"))


```

```{r detailedMap, echo=FALSE}

sub_region_out<-readRDS(file.path("data","final","int_results.rds"))

matched_points<-map(sub_region_out,~.x$bio_pnt_out)

keep_wu<-sapply(matched_points,function(x) nrow(x)>20)
keep_wu<-names(keep_wu)[keep_wu]
keep_wu<-str_extract(keep_wu,"^...")

matched_points2<-matched_points[str_extract(names(matched_points),"^...") %in% keep_wu]

matched_points2<-map2(matched_points2,names(matched_points2),~mutate(.x,WorkUnitName=.y)) %>% 
  bind_rows()

streams<-map(sub_regions,~map(.x,~.x$stream)) %>% 
    unlist(recursive = F) 

streams<-map2(streams,names(streams),~mutate(.x,WorkUnitName=.y)) %>% 
  bind_rows() %>% 
  filter(str_extract(WorkUnitName,"^...") %in% keep_wu)


tm_layout(main.title="Fish Sampling Locations Across AEC Subregions") +
  tm_shape(boundary %>% filter(str_extract(WorkUnitName,"^...") %in% keep_wu)) +
  tm_polygons() +
  tm_facets("WorkUnitName") +
  tm_shape(streams %>% filter(Strahler_Order>2)) +
  tm_lines(alpha = 0.5, col="darkblue") +
  tm_facets("WorkUnitName") +
  tm_shape(matched_points2) +
  tm_dots() +
  tm_facets("WorkUnitName") 
  

```

## Workflow

The data required for running models is too large to store on github, but can be downloaded from scripts included in this project.

Scripts in the R folder are numbered in the order they should be run. 

## Contribute

Fork the directory and copy locally. Feel free to send pull requests to merge changes into this main branch.


